---
title: "simpleRegress"
author: "Lacei Brown"
date: "Summer 2025"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

```{r}
bAGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(bAGE)
summary.lm(bAGE)
```

We want to compare the model with age to the other model with neck. Ordering by adjusted R square in the model with neck we got 0.7476 and for rhe age model it was 0.5073. With this we found out that neck is better thank age.  

```{r}
bHEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(bHEADLEN)
summary.lm(bHEADLEN)
```

After comparing the neck model, the AGE model and the HEADLEN model in order from biggest to smallest we got HEADLEN -0.8434, NECK -0.7476, and AGE -0.5073.

```{r}
bHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(bHEADWTH)
summary.lm(bHEADWTH)
```

After comparing the neck model, the AGE model, the HEADLEN model and the HEADWTH in order from biggest to smallest is HEADLEN -0.8434, NECK -0.7476, HEADWTH-o.5321 and AGE -0.5073.

```{r}
bCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(bCHEST)
summary.lm(bCHEST)
```



